Vulnerability Analysis of Container Images in Neuroimaging

### 2020-07-07 paper.tex
minor fixes

Clair (\url{https://github.com/quay/clair}) is an open-source and extensible
vulnerability scanner for Docker and appc container images, developed by
CoreOS (now Container Linux), a Linux distribution to deploy container
clusters. Clair has a client-server architecture, in which the server
maintains a database of vulnerabilities. We use Clair through Clair scanner
(\url{https://github.com/arminc/clair-scanner}), a tool to facilitate the testing
of container images against a local Clair server. Clair scanner scans the image,
prepares the list of vulnerabilities, compares that list against a
whitelist, and flags vulnerabilities that are not present in the whitelist.
In our experiments, we did not use a whitelist to filter scanning results in
order to make a fair comparison between scanners.
\tristan{What is this whitelist about, and which one did you use?}. Clair
scanner maintains a Docker image with the up-to-date vulnerability
database from a set of different sources listed in Figure~\ref{clair}.  \tristan{Please mention the vulnerability sources for all distributions}.
 We use Clair-scanner version 2.0.6, executed through
Docker image \texttt{arminc/clair-local-scan:v2.0.6}. For the vulnerability
database, we use Docker image \texttt{arminc/clair-db:latest}, last
updated on 2019-09-18.

@@

Singularity Tools (Stools)
are an extension of Clair scanner for Singularity images. Stools
directly scan the image files in a local installation of the Clair
scanner, to circumvent the Docker-specific requirements in the Clair API.
In our experiments, we use Singularity Tools version 3.2.1 through Docker
image
\texttt{vanessa/stools-clair:v3.2.1}, which is using another Docker
image \texttt{arminc/clair-db:latest} for referring to vulnerability feed.
We follow the steps mentioned in
\href{https://github.com/singularityhub/stools}{Stools Github} to scan
Singularity images. \tristan{which docker images did you use for clair and clair database?}

\subsection{Linux Releases and Vulnerability Databases}

\tristan{This section is still very confusing. I can rewrite it but would first need to
have a clear picture of vulnerability statuses across the tested Linux distributions. 
Are vulnerability statuses present only in Ubuntu, or are there statuses in other distributions too?
If so, what are the vulnerability statuses in the other distributions?
}
Through our experiments, we find that image scanners report vulnerabilities 
differently depending on the different Linux release types. So, it is 
important to provide an overview of these Linux releases types. 
BIDS app and Boutiques images in our experiments have diverse Linux
distributions, such as Ubuntu, CentOS, Debian, and Alpine. Further, a particular
distribution has different releases in general. To illustrate, we here examine Ubuntu as an example,
since it is heavily utilized in our experiments. 
\todo{Explain the situation with centos, update figure 3 accordingly.}
However, other distributions also have
similar release trend. Ubuntu releases
can be of two types: Regular or Long Term Support (LTS).
Regular release is supported only for 9 months, whereas LTS is supported for 5 years. Once LTS support
reaches End-Of-Life (EOL), Ubuntu decides whether to provide extended security support or not, which comes
in the form of Extended Security Maintenance (ESM) release.
ESM is a paid service that can be requested from Ubuntu to get security 
updates even after the EOL of a particular Ubuntu release. ESM and LTS releases of a particular Ubuntu distribution have different vulnerability
databases, because after EOL of a release, a vulnerability is fixed only in packages of ESM; however, it still
exists in LTS.

As an example, Ubuntu has ended support for some releases, such as 14.04 LTS,
17.04, and 17.10, as they reached EOL. However, Ubuntu decided to provide optional extended support for Ubuntu 14.04.
To use this extended support, users have to purchase ESM release for Ubuntu 14.04.
The images in our experiments have Ubuntu 14.04 LTS, so scanners should refer
to LTS release only while collecting data from the vulnerability databases.
More interestingly, scanners are able to refer to ESM release vulnerability database
even without purchase, however, this may lead to discrepancies in results.
%If scanners refer to ESM release for 
%vulnerability data, it may lead to discrepancies in results.
Further, if an image is using a Linux release where EOL has been reached that image may
turns to be vulnerable as it is not possible to retrieve any updates.
%Also, if ubuntu release with ended support is used in images, there are no chances of getting updates, hence
%resulting in vulnerable images.

Ubuntu maintains its vulnerability database in the Ubuntu Launchpad or the Ubuntu CVE Tracker. For a given CVE, there is an entry for all ubuntu releases.
Ubuntu releases are listed with the package name
in which vulnerability exists and are entitled a status
(\url{https://git.launchpad.net/ubuntu-cve-tracker/plain/README}), which is
encoded in the following form:
\newline \\
\noindent <release>\_<source-package>: <status> (<version/notes>) \\
\newline\\
For a given release, the status can be any one of the following:

@@

\textbf{False positives:} False positive reporting is also detected. For instance,
there are false positives in clair-scanner.
Some vulnerabilities are not linked to Ubuntu 14.04 LTS which is 
present inside the image, however, Clair-scanner have reported such vulnerabilities.
For instance, CVE-2019-1274 is an example of such vulnerabilities. 
%is showing because these vulnerabilities are present in Ubuntu 14.04 ESM and 
%Clair is only referring to Ubuntu 14.04 ESM release. 
%The reasons of Clair showing
%false positives and false negatives are unclear. We reported this issue on two Google groups of Clair i.e CoreOS Dev and
%clair-dev \href{https://groups.google.com/forum/\#!topic/clair-dev/8IqCCfd-EEc}{(https://groups.google.com/forum/\#!topic/clair-dev/8IqCCfd-EEc).}

>>>>

Clair (\url{https://github.com/quay/clair}) is an open-source and extensible
vulnerability scanner for Docker and appc container images, developed by
CoreOS (now Container Linux), a Linux distribution to deploy container
clusters. Clair has a client-server architecture, in which the server
scans Docker images layer by layer and maintains a database of vulnerabilities. We use Clair through Clair-scanner
(\url{https://github.com/arminc/clair-scanner}), a tool to facilitate the testing
of container images against a local Clair server. Clair-scanner scans the image,
prepares the list of vulnerabilities, compares that list against a
whitelist, and flags vulnerabilities that are not present in the whitelist.
In our experiments, we did not use a whitelist to filter scanning results in
order to make a fair comparison between scanners.
\tristan{What is this whitelist about, and which one did you use?}. Clair
scanner maintains a Docker image with the up-to-date vulnerability
database from a set of different sources listed in Figure~\ref{clair}.  \tristan{Please mention the vulnerability sources for all distributions}.
 We use Clair-scanner version 2.0.6, executed through
Docker image \texttt{arminc/clair-local-scan:v2.0.6}. For the vulnerability
database, we use Docker image \texttt{arminc/clair-db:latest}, last
updated on 2019-09-18.

@@

Singularity Tools (Stools)
are an extension of Clair scanner for Singularity images. Stools
exports Singularity image to \url{tar.gz} format, masquerading as a single layer docker image
to circumvent the Docker-specific requirements in the Clair API.
In our experiments, we use Singularity Tools version 3.2.1 through Docker
image
\texttt{vanessa/stools-clair:v3.2.1}, which is using another Docker
image \texttt{arminc/clair-db:latest} for referring to vulnerability feed.
We follow the steps mentioned in
\href{https://github.com/singularityhub/stools}{Stools Github} to scan
Singularity images. \tristan{which docker images did you use for clair and clair database?}

\subsection{Ubuntu Releases and its Vulnerability Database}

\tristan{This section is still very confusing. I can rewrite it but would first need to
have a clear picture of vulnerability statuses across the tested Linux distributions. 
Are vulnerability statuses present only in Ubuntu, or are there statuses in other distributions too?
If so, what are the vulnerability statuses in the other distributions?
}
BIDS app and Boutiques images in our experiments have diverse Linux
distributions, such as Ubuntu, CentOS, Debian, and Alpine. A particular
distribution has different releases in general. Regarding Ubuntu, it has different
types of releases e.g. Regular, Long Term Support (LTS), and Extended Security Maintenance (ESM) release.
Through our experiments, we find that image scanners report Ubuntu vulnerabilities 
differently depending on the different Ubuntu release types. It is 
important to provide an overview of these Ubuntu releases types, 
since Ubuntu is heavily utilized in our experiments. 
\todo{Explain the situation with centos, update figure 3 accordingly.}
Ubuntu's Regular release is supported only for 9 months, whereas LTS is supported for 5 years. Once LTS support
reaches End-Of-Life (EOL), Ubuntu decides whether to provide extended security support or not, which comes
in the form of ESM release.
ESM is a paid service that can be requested from Ubuntu to get security 
updates even after the EOL of a particular Ubuntu release. ESM and LTS releases of a particular Ubuntu version have different vulnerability
databases, because after EOL of a release, a vulnerability is fixed only in packages of ESM; however, it still
exists in LTS.

As an example, Ubuntu has ended support for some releases, such as 14.04 LTS,
17.04, and 17.10, as they reached EOL. However, Ubuntu decided to provide optional extended support for Ubuntu 14.04.
To use this extended support, users have to purchase Ubuntu 14.04 ESM release.
The images in our experiments have Ubuntu 14.04 LTS, so scanners should refer
to LTS release only while collecting data from the vulnerability databases.
More interestingly, scanners are able to refer to ESM release vulnerability database
even without purchase, however, this may lead to discrepancies in results.
%If scanners refer to ESM release for 
%vulnerability data, it may lead to discrepancies in results.
Further, if an image is using a Linux release where EOL has been reached that image may
turns to be vulnerable as it is not possible to retrieve any updates.
%Also, if ubuntu release with ended support is used in images, there are no chances of getting updates, hence
%resulting in vulnerable images.

Ubuntu maintains its vulnerability database in the Ubuntu CVE Tracker. For a given CVE, there is an entry for all ubuntu releases.
Ubuntu releases are listed with the package name
in which vulnerability exists and are entitled a status
(\url{https://git.launchpad.net/ubuntu-cve-tracker/plain/README}), which is
encoded in the following form:
\newline \\
\noindent <release>\_<source-package>: <status> (<version/notes>) \\
\newline\\
For a given release, the status can be any one of the following:

@@

\textbf{False positives:} False positive reporting is also detected. For instance,
there are false positives in clair-scanner.
Some vulnerabilities are not linked to Ubuntu 14.04 LTS which is 
present inside the image, however, Clair-scanner have reported such vulnerabilities.
For instance, CVE-2019-1274 is an example of such vulnerabilities. 
%is showing because these vulnerabilities are present in Ubuntu 14.04 ESM and 
%Clair is only referring to Ubuntu 14.04 ESM release. 
The reasons of Clair showing
false positives and false negatives are unclear. We reported this issue on two Google groups of Clair i.e CoreOS Dev and
clair-dev \href{https://groups.google.com/forum/\#!topic/clair-dev/8IqCCfd-EEc}{(https://groups.google.com/forum/\#!topic/clair-dev/8IqCCfd-EEc).}
